---
title: "D4 Module 1 - Dirty Data Project"
author: "Steve Maddison"
date: "24/07/2020"
output:
  html_document:
    toc: true
    toc_float: true
---

# Load libraries

```{r}
library(tidyverse)
library(dplyr)
library(janitor)
library(here)
library(readr)
library(readxl)
```

# Read cleaned data file

```{r}
# gender and country need to be defined col_type
candy <- read_csv("../clean_data/boing_boing_candy_clean.csv",
                  col_types = cols(gender = col_character(), country = col_character()))
```

# Review data

```{r}
dim(candy)
```

```{r}
glimpse(candy)
```

# TASK QUESTIONS
 
## Q1: What is the total number of candy ratings given across the three years. (number of candy ratings, not number of raters. Don’t count missing values)

```{r}
total_candy_ratings <- candy %>%
  select(rating) %>%
  filter(rating == "joy" | rating == "despair" | rating == "meh") %>%
  summarise(total_ratings = n())
total_candy_ratings
```

## Q2: What was the average age of people who are going out trick or treating and the average age of people not going trick or treating?

```{r}
average_age_going_out <- candy %>%
  select(age, trick_or_treating) %>%
  filter(trick_or_treating == "yes", age != "NA") %>%
  summarise(average_age_going_out = mean(age))
average_age_going_out

average_age_not_going_out <- candy %>%
  select(age, trick_or_treating) %>%
  filter(trick_or_treating == "no", age != "NA") %>%
  summarise(average_age_not_going_out = mean(age))
average_age_not_going_out
```

## Q3: For each of joy, despair and meh, which candy bar received the most of these ratings?

```{r}
candy_with_most_joy_rating <- candy %>%
  select(rating, candy) %>%
  filter(rating == "joy") %>%
  group_by(candy) %>%
  summarise(candy_with_most_joy_rating = n()) %>%
  arrange(desc(candy_with_most_joy_rating)) %>%
  head(1)
candy_with_most_joy_rating

candy_with_most_despair_rating <- candy %>%
  select(rating, candy) %>%
  filter(rating == "despair") %>%
  group_by(candy) %>%
  summarise(candy_with_most_despair_rating = n()) %>%
  arrange(desc(candy_with_most_despair_rating)) %>%
  head(1)
candy_with_most_despair_rating

candy_with_most_meh_rating <- candy %>%
  select(rating, candy) %>%
  filter(rating == "meh") %>%
  group_by(candy) %>%
  summarise(candy_with_most_meh_rating = n()) %>%
  arrange(desc(candy_with_most_meh_rating)) %>%
  head(1)
candy_with_most_meh_rating
```

## Q4: How many people rated Starburst as despair?

```{r}
starburst_despair_rating <- candy %>%
  select(candy, rating) %>%
  filter(candy == "starburst", rating == "despair") %>%
  group_by(candy) %>%
  summarise(starburst_despair_rating = n())
starburst_despair_rating
```

## For the next three questions, count despair as -1, joy as +1 and meh as 0.

## Q5: What was the most popular candy bar by this rating system for each gender in the dataset?

```{r}
unique(candy$gender)
```

```{r}
popular_candy_by_gender_male <- candy %>%
  select(gender, candy, rating) %>%
  filter(gender == "male" & rating == "joy")
  #mutate(new_rating = count(rating) * 1)
popular_candy_by_gender_male
```


