---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
library(dplyr)
library(janitor)
library(here)
library(readr)
```

```{r}
# read clean data from CSV file into dataframe
decathlon <- read_csv(here("clean_data/decathlon_clean.csv"))
decathlon
```

Q1: Who had the longest long jump seen in the data?

```{r}
# filter data frame for long_jump rows only, then order results
# descending and output the top row
longest_long_jump <- decathlon %>%
  filter(event == "long_jump") %>%
  arrange(desc(result)) %>%
  head(1)
longest_long_jump
```

Q2: What was the average 100m time in each competition?

```{r}
# select the applicable columns, filter for 100m rows only, group by
# competition and output the average of the results 
avg_100m <- decathlon %>%
  select("competition", "event", "result") %>%
  filter(event == "x100m") %>%
  group_by(competition) %>%
  summarise(average_result = mean(result))
avg_100m
```

Q3: Who had the highest total points across both competitions?

```{r}
# select the applicable columns, group by name and points, summarise
# the results then order results descending and output the top row
highest_points <- decathlon %>%
  select("name", "points") %>%
  group_by(name, points) %>%
  summarise() %>%
  arrange(desc(points)) %>%
  head(1)
highest_points
```

Q4: What was the shot-put scores for the top three competitors in each competition?

# Note: I'd prefer to code this answer using group_by on the competition column
# rather than hard-code the competition names (and to avoid code duplication)
# but I couldn't get the results to order correctly using group_by or output
# the top row of each grouped competition

```{r}
# select the applicable rows, filter for competition and event then order
# the results descending then output the top 3 rows
top_3_shot_put_decastar <- decathlon %>%
  select("name", "competition", "result", "event") %>%
  filter(competition == "Decastar", event == "shot_put") %>%
  arrange(desc(result)) %>%
  head(3)
top_3_shot_put_decastar
```

```{r}
# select the applicable rows, filter for competition and event then order
# the results descending then output the top 3 rows
top_3_shot_put_olympicg <- decathlon %>%
  select("name", "competition", "result", "event") %>%
  filter(competition == "OlympicG", event == "shot_put") %>%
  arrange(desc(result)) %>%
  head(3)
top_3_shot_put_olympicg
```

Q5: What was the average points for competitors who ran the 400m in less than 50 seconds vs. those than ran 400m in more than 50 seconds?

# Note: I would prefer this code to be a single chunk rather than 2 separate
# chunks to avoid code duplication

```{r}
# select applicable columns, filter by event and result time, groups by event
# and output the summary of the average result
avg_400m_less_50s <- decathlon %>%
  select("result", "event") %>%
  filter(event == "x400m", result < 50) %>%
  group_by(event) %>%
  summarise(average_result = mean(result))
avg_400m_less_50s
```

```{r}
# select applicable columns, filter by event and result time, groups by event
# and output the summary of the average result
avg_400m_more_50s <- decathlon %>%
  select("result", "event") %>%
  filter(event == "x400m", result > 50) %>%
  group_by(event) %>%
  summarise(average_result = mean(result))
avg_400m_more_50s
```
